---
title: '`Practice Quiz` Regression Models'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
  html_document:
    self_contained: true
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  author: 'Anderson H Uyekita'
  course: 'Regression Models'
  course_url: 'https://www.coursera.org/learn/regression-models'
  specialization: 'Data Science: Foundations using R Specialization'
  specialization_url: 'https://www.coursera.org/specializations/data-science-foundations-r'
  instructor: 'Brian Caffo'
  course_start: !r lubridate::ymd('2022/07/23', tz = 'America/Sao_Paulo')
  course_finish: !r lubridate::ymd('2022/07/25', tz = 'America/Sao_Paulo')
  week: "Week 3"
  slug: 'Week%203'
  rpubs: 'https://rpubs.com/AndersonUyekita/course-project_regression-models'
  project_repo: 'https://github.com/AndersonUyekita/regression-models_course-project'
  codebook: 'https://github.com/AndersonUyekita/regression-models_course-project/blob/master/codebook.md'
  instructions: 'https://github.com/AndersonUyekita/regression-models_course-project/blob/master/instructions.md'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo=TRUE,cache=TRUE,message=FALSE,warning=FALSE)
library(rmarkdown)
library(matahari)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F30E; Rpubs: [Interactive Document](%s)", params$rpubs)}else{sprintf(fmt = "* &#x1F4E6; Github Repository: [Static Document](%s){target='_blank' rel='noopener'}", params$project_repo)}`

--------------------------------------------------------------------------------

## Brief Analysis about Manual and Automatic Transmissions

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/AndersonUyekita/regression-models_practice_quiz/main?urlpath=rstudio)

```{r record_start,echo=FALSE}
# Start the command recording.
matahari::dance_start(value = FALSE, contents = FALSE)
```


## Synopsis


## 1. Introduction

The Optional Quiz Assignment wants to analyze the relationship between income and major categories. This study should be performed using the `college` dataset from the `collegeIncome` library.

## 2. Objectives

This Practice Quiz aims to answer the following question:

> Based on your analysis, would you conclude that there is a significant association between college major category and income?

- [ ] Yes
- [ ] No

## 3. Requeriments

It is necessary to use the following packages to perform this experiment.

```{r libraries,echo=TRUE}
# Loading packages
library(collegeIncome)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(explore)
library(kableExtra)
library(DT)
library(PerformanceAnalytics)
library(DiagrammeR)
library(GGally)
```

### 3.1. Reproducibility

If you want to reproduce it, please, fork the experiment repository hosted on Github.

* [Github Repository](https://github.com/AndersonUyekita/regression-models_practice_quiz)


### 3.2. Loading Data

Following the practice quiz instructions, I have used the `college` dataset from the `collegeIncome` package.

```{r loading_data,echo=TRUE}
# Loading college data to environment.
data("college")

# Creating a copy of college data.
df_college <- college
```

### 3.3. Codebook

From the assignment instructions:

* rank: Rank by median earnings
* major_code: Major code
* major: Major description
* major_category: Category of major
* total: Total number of people with major
* sample_size: Sample size of full-time, year-round individuals used for income/earnings estimates: p25th, median, p75th
* p25th: 25th percentile of earnings
* median: Median earnings of full-time, year-round workers
* p75th: 75th percentile of earnings
* perc_men: % men with major (out of total)
* perc_women: % women with major (out of total)
* perc_employed: % employed (out of total)
* perc_employed_fulltime: % employed 35 hours or more (out of employed)
* perc_employed_parttime: % employed less than 35 hours (out of employed)
* perc_employed_fulltime_yearround: % employed at least 50 weeks and at least 35 hours (out of employed and full-time)
* perc_unemployed: % unemployed (out of employed)
* perc_college_jobs: % with job requiring a college degree (out of employed)
* perc_non_college_jobs: % with job not requiring a college degree (out of employed)
* perc_low_wage_jobs: % in low-wage service jobs (out of total)

## 4. Exploratory Analysis

### 4.1. Dataset Dimensions

The `college` dataset has 173 observations and 19 variables.

```{r dimensions,echo=TRUE}
# Checking the number of observations and variables.
dim(df_college)
```

### 4.2. Head

The first 10 rows of the dataset:

```{r head,echo=FALSE}
# Printing the first five rows.
if (knitr::is_html_output()) {
    # RPUB
    df_college %>% DT::datatable(options = list(searching = FALSE,scrollX = T))
} else {
    # Github
    head(df_college, 10) %>% kbl()
}
```

### 4.3. Tail

The last three rows of the dataset:

```{r tail,echo=FALSE}
# Printing the first five rows.
if (knitr::is_html_output()) {
    # RPUB
    df_college %>% DT::datatable(options = list(searching = FALSE, scrollX = TRUE),
                                 callback = JS('table.page("last").draw(false);'))
} else {
    # Github
    tail(df_college, 10) %>% kbl()
}
```

### 4.4. Structure

Let's check the variables' types.

```{r str,echo=FALSE}
# Checking the data types.
df_college %>% str()
```

One observes there are problems with variables types:

* `major`: Should convert it into a category;
* `major_category`: Should convert it into a category.

Also, some variables are not helpful for the analysis:

* `rank`: There is no info about how this rank was calculated.
* `major_code`: The major code is a primary key, so there is no reason to use it in a deeper analysis.

### 4.5. Summary

To confirm the presence of NA observations, categorical variables as characters, and other problems. Let's print the `summary()`.

```{r summary,echo=FALSE}
# Printing the summary.
summary(df_college)
```

### 4.6. Data Visualization

#### 4.6.1. Numeric Variables

The following graph shows a density plot of each numeric variable.

```{r explore_numeric,echo=FALSE,fig.height=18}
# Checking the data types.
df_college %>% 
    dplyr::select(-c(rank, major_code, major, major_category)) %>%
    explore_all()
```

Unfortunately, some variables have `NA` values, which will be required to clean them. So, the following observations have one or more `NA`, `Inf`, or invalid content.

```{r observations_issues,echo=FALSE}
# Printing observations with problems.
if (knitr::is_html_output()) {
    # RPUB
    df_college %>%
        dplyr::filter(if_any(everything(), is.na)) %>%
        DT::datatable(options = list(searching = FALSE, scrollX = TRUE),
                                 callback = JS('table.page("last").draw(false);'))
} else {
    # Github
    df_college %>%
        dplyr::filter(if_any(everything(), is.na)) %>%
        kbl()
}
```

`Industrial And Manufacturing Engineering` and `Computer And Information Systems` majors contain invalid values.

#### 4.6.2. Categorical Variables

There is no way to plot a visible graph to `major` because there are 173 categories in this variable. For this reason, I will not plot any graph for it.

As expected, the `major` variables have 173 unique values, meaning each row corresponds to a unique major. However, remember that I have not inspected each `major` name, so I can not ensure if it contains typos or the same major with different notations.

```{r explore_categorical_major_category,echo=FALSE,fig.height=3}
# Checking the data types.
df_college %>% 
    dplyr::select(major, major_category) %>%
    dplyr::mutate(major = factor(major), major_category = factor(major_category)) %>%
    explore(major_category)
```

The `major_category` variable have 16 categories. Table 1 summarizes all majors in respect of `total` and `sample_size.`

```{r major_category_summary,echo=FALSE}
# Counting the number of major for each major_category.
df_college %>%
    group_by(major_category) %>%
    summarize(number_major = n(), total = sum(total), sample_size = sum(sample_size)) %>%
    arrange(desc(number_major)) %>%
    kbl() %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = TRUE)
```

Highlights:

* `Interdisciplinary` corresponds to 0.6% of total major courses. In absolute terms, this category has only one major course (`Multi/Interdisciplinary Studies`);
* Due to the `size_sample` and `total` of the `Interdisciplinary` category, it is convenient to remove it.
* `Engineering` is the `major_category` with the most number of majors, and;
* `Business` is the `major_category` with the most number of students..

#### 4.6.3. Variables Correlation

I will show the scatter plot with histogram and correlation between each variable.

```{r data_correlation,echo=FALSE,fig.height=18, fig.width=18}
# Printing the correlation, histogram and scatter plot.
df_college %>% 
    dplyr::mutate(
        major_code = factor(major_code),
        major_category = factor(major_category)
    ) %>%
    dplyr::select(-c(rank, major_code, major, major_category, p25th, p75th)) %>%
    drop_na() %>%
    chart.Correlation(histogram = TRUE, pch = 19)
```

Highlights:

* `total` and `sample_size` have a high correlation because the greater the number of people with that major, the greater the sample_size. It is necessary to drop one out of two;
* `perc_women` and `perc_men` have a perfect correlation, which is expected because they are complementary;
* `perc_employed` and `perc_unemployed` also have a perfect correlation, which is expected because they are complementary;
* `perc_employed_fulltime` and `perc_employed_parttime` have high correlation. Given that you are employed, you only have two options, full-time or part-time, so those variables are also complementary;
* `perc_college_jobs` and `perc_non_college_jobs` have high correlation. Given that you are employed, you only have two options to the job type, college or non-college, so those variables are also complementary, and;
* `perc_low_wage_jobs` is positive correlated with `perc_non_college_jobs` and negative correlated with `perc_college_jobs` because it is expected that college jobs has greater payloads in comparison to non-college jobs, and.

## 5. Model Selection

Based on section 4, Exploratory Data Analysis, there are several variables that I cannot use in the Model Selection due to a high correlation. For this reason, I will drop the following variables:

* `rank`;
* `major_code`;
* `major`;
* `sample_size`;
* `perc_men`;
* `perc_unemployed`;
* `perc_employed_fulltime`;
* `perc_college_jobs`;
* `p25th`, and;
* `p75th`.

The figure below shows the variable relationship.

```{r digraph,echo=FALSE}
# Creating a diagram to explain the job condition.
mermaid("
graph LR
  X(Begin)-->A[total]
  A-->B{Am I employed?}
  B-->G[perc_employed]
  G-->C{Is it full-time or part-time job?}
  C-->E[perc_employed_fulltime]
  E-->M{Full-time year arround?}
  M-->N[perc_employed_fulltime_yearround]
  C-->F[perc_employed_parttime]
  G-->H{Is it college or non-college job?}
  H-->J[perc_college_jobs]
  H-->L[perc_non_college_jobs]
  A-->I{Is it a low wage jobs?}
  I-->K[perc_low_wage_jobs]
  B-->D[perc_unemployed]
  A-->O{Gender?}
  O-->P[perc_men]
  O-->Q[perc_women]
")
```

### 5.1. Data Manipulation

To convert the `college` dataset into a tidy dataset, it is mandatory to:

* Subset the college dataset by selecting the variables: `major_category`, `total`, `perc_women`, `median`, `perc_employed`, `perc_employed_parttime`, `perc_employed_fulltime_yearround`, `perc_non_college_jobs`, and `perc_low_wage_jobs`;
* Converting misclassified variables as characters into factors;
* Removing `NA` observations, and;
* Removing non-representative categories.

```{r data_cleaning,echo=FALSE}
# Data Cleaning.
df_tidy <- df_college %>%
    
    # Subsetting
    dplyr::select(-c(rank, major_code, major,sample_size, perc_men, perc_unemployed, perc_employed_fulltime, perc_college_jobs)) %>%
    
    # Converting character variables into factor.
    dplyr::mutate(major_category = factor(major_category)) %>%
    
    # Dropping NA observations.
    tidyr::drop_na() %>%
    
    # Removing non-representative categories.
    dplyr::filter(major_category != "Interdisciplinary")
```

The tidy dataset has 9 variables and 170 observations.

```{r}
# Checking the tidy dataset dimensions.
dim(df_tidy)
```

I will count the NA values using the `is.na()` function to ensure I have eliminated them.

```{r testing_tidy_dataset}
# Testing
sum(is.na(df_tidy))
```

It is zero, which means there are no `NA` values.

### 5.2. 25th, 50th or 75th Percentile

Regardless of the dependent variable used as output, the three options have a high correlation. So, the figure below will show the correlation matrix between the 25th, 50th, and 75th percentile.

```{r outcome_correlation,echo=FALSE}
# Printing a correlation plot with 25th, 50th, and 75th percentile.
df_tidy %>% 
    dplyr::select(c(p25th, median, p75th)) %>%
    chart.Correlation(histogram = TRUE, pch = 19)
```

Highlight:

* `p25th`, `median`, and `p75th` are strongly positively correlated.

Any of those used as the dependent variables will perform similar outcomes.

### 5.3. Gender Income Gap

According to the Bureau of Labor Statistics, the income gap between men and women is around 62 USD per week. Let's divide the women penetration in a given major category into 4 levels:

* Category A: `perc_women` below 25%;
* Category B: `perc_women` between 25% and 50%;
* Category C: `perc_women` between 50% and 75%, and;
* Category D: `perc_women` above 25%.

```{r outcome_correlation_by_category,echo=FALSE}
# Printing a correlation plot with 25th, 50th, and 75th percentile.
df_tidy %>% 
    
    # Selecting income variables and perc_women.
    dplyr::select(c(p25th, median, p75th, perc_women)) %>%
    
    
    dplyr::mutate(a = factor(
        case_when(
            perc_women <  0.25 ~ "A",
            perc_women >= 0.25 & perc_women < 0.50 ~ "B",
            perc_women >= 0.50 & perc_women < 0.75 ~ "C",
            perc_women >= 0.75 ~ "D"
        )
    )) %>%
    
        ggpairs(columns = 1:4, aes(color = a, alpha = 0.5))
```

The above graph shows no difference between high women penetration in the yearly income, which is a bit counter-intuitive due to the well-known gap between men's and women's wages. Moreover, all density curves are barely the same or with minor changes.

### 5.4. Linear Regression

Considering income is directly related to how much time you spend working, the `perc_employed_fulltime_yearround` will play a key role. Also, people who does not work in college jobs (`perc_non_college_jobs`) will have a lower wage, which is also related to `perc_low_wage_jobs`. From the section 5.3. Gender Income Gap, I will drop the gender variable (`perc_women`).


$$median = \beta_0 + \beta_1 \cdot major\_category + \beta_2 \cdot perc\_employed\_fulltime\_yearround + \beta_3 \cdot perc\_non\_college\_jobs + \beta_4 \cdot perc\_low\_wage\_jobs$$

```{r,echo=FALSE}
# Fitting a model
fit <- lm(data = df_tidy %>% select(-c(p25th, p75th)), formula = median ~ major_category + perc_employed_fulltime_yearround + perc_non_college_jobs + perc_low_wage_jobs - 1)

# Printing the coefficients.
summary(fit)$coeff %>% round(4)
```

From the above results of the `lm()` function, there is no statistical evidence that `perc_employed_fulltime_yearround`, `perc_non_college_jobs`, `perc_low_wage_jobs`, and the dummy variable `major_category` affect the income. Almost all p-value failed to reject the $H_0$ hypothesis.

## 6. Results

The answer to the posed question is: There is **no** significant association between college and major category and income.

```{r record_save,echo=FALSE}
# Save and stop the command recording.
matahari::dance_save("./college_major_analysis.rds")
matahari::dance_stop()
```